### 2020-12-30 BMC-Journal of Biomedical Semantics/submission 0/bmc_template/results.tex
add new section pages

\section*{Results}
\label{results}
Results goes here ....