### 2021-01-12 BMC-Journal of Biomedical Semantics/submission 0/bmc_template/experimentalsetup.tex
method 1

\section*{Experimental Setup}
\label{experimentalSetup}