### 2021-06-02 BMC-Journal of Biomedical Semantics/submission 0/bmc_template/methodology.tex

\section*{Methodology}
\label{methodology}
Healthcare interoperability, with a focus on non-standard compliant medical schema, is dependent on the generation and validation of schema maps, as discussed above. To this end, the creation of a cohesive workflow is of utmost importance. In our earlier work \cite{Satti2020} we used maximum sequence identification and suffix tress based matching for syntactic matching of two distinct data schemas. This was followed by semantic concept enrichment and subsequently concept matching, for creating rules in the form of schema maps. The simplified mapping functions, thereby created, provided a simple methodology for converting semi-structured medical data, into an interpretable, model form.
In our current methodology we have utilized state-of-the-art natural language processing (NLP) techniques to extract the schema mapping rules from semi-structured data schemas. This methodology is based on identifying similarity between vector representations of two attributes, belonging to different medical schemas. 
\textbf{Traditional NLP techniques such as WordNet are able to convert a word into an embedded vector, while Bidirectional Encoder Representations from Transformers (BERT) extracts an embedded vector from a sentence.} However, the terms forming the attribute names are bigger than a word (combination of multiple words) and smaller than a sentence. In order to resolve this problem, we extracted the set of suffixes from the terms forming the attribute names. The bidirectional nature of BERT, allows the creation of contextual embedded vectors, where each target word is affected by its neighboring words. Hence to convert the set of suffixes into a sentence, we collected the set of concepts corresponding to each suffix, from Unified Medical Language System (UMLS). This operation has two effects, firstly it is used to remove any suffix, which does not have a corresponding concept and secondly the extracted concepts are used to add context to each suffix and produce a contextual sentence. 
The following subsections provide the practical details for our methodology, from schema acquisition, to attribute name expansion, and finally schema map generation.

\subsection*{Schema acquisition}
\label{schema_acq}
In the first step of our semantic reconciliation methodology, we simulate medical data acquisition from five distinct EMR storage systems ($S$). These include patient reports from OpenEMR ($s_1$), 100,000 patient records from EMRBOTS ($s_2$) \cite{kartoun2016methodology}, custom database design by Pan et. al($s_3$) for supporting regional clinics and health care centers in China \cite{pan2016design}, clinical knowledge discovery tool MedTAKMI-CDI ($s_4$) \cite{inokuchi2007medtakmi}, and our custom implementation ($s_5$). Each of these medical systems follow the relational database design, with logical entities, such as demographics, diagnosis, medicine or others, placed into tables which can be further linked to one or more tables. While the database design implemented by each of these systems, fulfills the need of their respective information processing applications, the lack of interoperability, in terms of identifying similar attributes or exchanging the medical data is very much evident here. 
A similar notion of data heterogeneity, in terms of medical data schema, is evident across the healthcare domain. This is caused by various factors, including the lack of one all-encompassing, and universally applicable terminological standard and different normalization level for representing attributes.\\
In the former case, while SNOMED-CT provides a mechanism for identifying the standard codes for clinical terms and LOINC can be used for laboratory related terms, most attribute names are created based on the gut feeling of the database designer. Additionally, while these codes can be used to represent data instances, the data schema, achieves no benefit from the same.  Consider the terms ``name" and ``patientName", which refer to the same attribute of the patient entity.  However, since there is no standard way to represent this attribute, both are considered correct ($s_1$ and $s_3$ use the former representation, while $s_2$ and $s_5$ use the latter).\\
In the later case, differences in normalization cause semantic differences, due to which some data could be available in one schema but absent in others, such as OpenEMR demographics identifying the patient's residential location using specific attributes like ``Address", ``City", ``State", ``Postal Code", ``Country", and others. Similarly, ``EncounterDate" from $s_5$ is semantically similar to ``BeginDate" of ``openemr\_MedicalProblems" table in $s_1$, ``AdmissionStartDate" of ``LabsCorePopulatedTable" in $s_2$, 
``time" in ``Diagnosis" table of $s_3$, and  ``dateOfAdmission" in ``Diagnosis" and ``CareHistory" tables of $s_4$.
Finally, $s_1$ and $s_3$ have separate tables containing the medicinal prescription details, however the same details are unavailable in $s_2$, $s_4$, $s_5$. Once again, this is not an incorrect behavior since this information, might not be a part of the context or the requirements for these EMR storage systems.
In fact, the change in context of the EMR storage system from the initial time of development to a later stage of collaborative processing systems, is the main cause of heterogeneity. In order to provide an interoperable solution, it is therefore necessary to enhance the semantics of each EMR attribute by its contextually equivalent sentence.

\subsection*{From Attribute to Sentence}
\label{schema_prepro}
In order to process the EMR schema set $S$ and produce a set of corresponding semantically enriched sentences, we use the methodology shown in \textbf{figAlgorithm}.
First we use the data representation $s_i$, generated through the process explained above (\ref{schema_acq}) to collect the various medical fragments in memory. We then iterate over these fragments, building a set of attributes, distinguished by their name, schema's name, table's name, schema's version, source, and recorded data. This entails that ``PatientID" from each of the four tables in $s_2$, and ``patientID" from five tables in $s_4$, would result into nine attributes (assuming, as in the current case, of no differences in versions of these systems).
For each attribute, we then generate the suffix array, which provides all possible substring representations contained within the attribute name. In order to generate the set of suffixes, we employ three strategies, forward suffix generation, whereby for a word $w$ of length $n$, $n-1$ suffixes of size 2 to $n-1$ are produced, backward suffix generation, to produce $n-1$ suffixes in reverse order with size $n-1$ to 2, and regular expression based suffix generation, which splits each word on, change of case, special characters(such as -, \_, !, and others), and numbers. An example of this suffix generation process is shown in \textbf{figSuffixAttribute}. 
Here, the attribute ``DOB" in $s_1$, ``PatientDateOfBirth" appears in $s_2$, ``birthday" appears in $s_3$, ``DateOfBirth" in $s_5$, and ``birthDate" in $s_4$. These attributes refer to the same attribute of the patient entity, however, with the suffix generation process, our aim is to identify all words within these strings. While the forward and backward generation process, generates all possible sequences of characters as suffix, the regex based method, only identifies the changes in the string. In this way a large list of suffixes is generated, which is combined using a ``TreeSet" data structure of Java, which internally sorts this list as well.

Similarly the attribute, ``dateOfAdmission" appears in $s_4$, while ``AdmissionStartDate". ``diseaseNameOnAdmission", and ``AdmissionEndDate" appear in $s_2$. As evident within these strings, while they refer to semantically different entity attributes, syntactically they contain similar elements. The suffixes generated from these strings are shown in \textbf{figSuffixAttribute (b)}. The suffix generation process, identified thus far, is only able to generate syntactic suffixes, producing many incoherent and unrelated suffixes. In order to counter this problem, and to limit the list of suffixes within the domain, we then query UMLS, with exact search strategy, looking for the existence of any concepts, against each suffix. In case, no semantic concept is found for a particular suffix, it is removed from the final Suffix Array. On the other hand, if atleast one semantic concept is found against the queried suffix, it is retained. Meanwhile the process continues for the next attribute, then the next table, and finally the next system, till no further processing is possible. 
The set of suffixes and their corresponding concepts are then used to build the sentence, where by each concept, corresponding to a suffix is appended next to the suffix. An example of the resultant sentence for the attribute ``DateOfAdmission" is shown as follows:\\

\framebox{\parbox{\dimexpr\linewidth-2\fboxsep-2\fboxrule}{\itshape%

		``Date Value type - Date date allergenic extract Date in time Data types - Date Date Fruit;Of SPI1 wt Allele SPI1 gene TAF1 wt Allele BRIP1 gene Within Degrees fahrenheit Oral contraception BRIP1 wt Allele;Da Displacement of abomasum dalton Anterior descending branch of left coronary artery deca units cytarabine/daunorubicin protocol Dai Chinese Asymptomatic diagnosis of Drug Accountability Domain;ion Iontophoresis Route of Drug Administration Ions;on SPARC wt Allele Osteonectin SPARC gene On (qualifier value) Upon - dosing instruction fragment;Admission Admission activity Hospital admission;Dat SLC6A3 gene SLC6A3 wt Allele dopamine transporter Direct Coombs test SLC6A3 protein, human Test Date cytarabine/daunorubicin/thioguanine Alzheimer's Disease;mission Religious Missions;"
}}\\

Here the various suffixes and their concepts are seperated by the symbol ``;", however together they form one sentence, for which an embedded vector is generated.

\subsection*{Schema Map generation}
\label{schema_map_gen}

Schema Maps, provide an interoperable bridge between two medical systems ($s_i \wedge s_j$), by identifying the semantic relationship between their participating attributes. This identification is based on the similarity between the embedded vectors, of the semantically enriched sentences, corresponding to each EMR attribute. While the embedded vectors can be generated using any methodology (we tested 11 methodologies, with WordNet and 10 models of BERT further detailed in section \ref{experimentalSetup}), the large/STSb version of the Siamese BERT-Networks \cite{reimers-2019-sentence-bert}, commonly known as \textbf{stsb-roberta-large} provides the best results. The pair of embedded vectors thus produced are then used to calculate cosine similarity, which is based on the inverse cosine distance between them. For our classification, we used the raw results (unnormalized) of cosine similarity, which produces a score between -1, and 1. Cosine similarity score of 0 indicates orthogonal relationship between the two vectors, which in our scenario indicates that the two sentences, and by extension their attributes are not related to each other. -1 indicates, inverse relationship between the attributes, while 1 indicates the two attributes are very much the same.
For producing our schema maps, we are interested in three types of relationships, ``equal" (the two attributes are same), ``related" (the two attributes are related to each other), and ``unrelated" (no relationship between the attributes). In order to classify the similarity results, into one of these three classes, we then calculated the best thresholds, using Matthews Correlation coefficient (MCC) \cite{chicco2020advantages} for classifying each instance as ``equal", ``related", and ``unrelated". Finally on a test dataset we evaluated our multi-class classification approach, to identify the relationships between each pair of attributes.