### 2021-06-03 BMC-Journal of Biomedical Semantics/submission 0/bmc_template/experimentalsetup.tex
update

\section*{Experimental Setup}
\label{experimentalSetup}
In our earlier work \cite{Satti2020} $s_1$, $s_2$, and $s_5$ were used to generate over 115 million patient records, which are converted into a semi-structured form and stored in Hadoop Distributed File System (HDFS). We extended the same setup to create an additional 100,000 records, for 1000 patients with 3 medical fragments for $s_1$, $s_2$, and $s_4$, and 97 randomly selected and generated medical fragments amongst $s_1$, $s_2$, $s_3$, $s_4$ and $s_5$. These fragments, follow various design elements, supporting a variety of valid relational storage architectures. Such as, $s_1$, $s_2$ and $s_4$ are represented by creating a separate medical fragment for each participating table, $s_3$ utilizes its medical fragment to generate a linked record (from a linked object graph), where by the attributes can refer to other objects besides the elements of $t$, mimicking the application of explicit foreign keys, and $s_5$ is a flat table structure. The code to generate this data set is available at ``uhp\_map\_generation"\footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhpr\_storage}. This application produces three custom formatted files, containing an index for patients, an index for their medical fragments, and the medical fragment, corresponding to the EMR data. Using the medical fragments file, we then generate the semantically enriched attribute \footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhp\_map\_generation}, which contains the suffixes and their concepts corresponding to each EMR data attribute. The resulting set of enriched attributes are temporarily stored in a ``json" file, which is then read by the same application to partially generate the schema maps. This process, is used to create 20,349 distinct pairs of attributes, across $s$. Each pair also contains the ``relationshipList", which stores the results of fuzzy string matching\cite{FuzzyWuzzy} \footnote{Java Library: https://github.com/xdrop/fuzzywuzzy} between the attribute names. 
The ``json" file thus produced, is then used by a python script to generate the semantically enriched sentences and their embedded vectors using Word2Vec, and 10 pre-trained BERT NLI models \cite{reimers-2019-sentence-bert}. The BERT models include 'bert-base-nli-stsb-mean-tokens', 'bert-large-nli-stsb-mean-tokens', 'roberta-base-nli-stsb-mean-tokens', 'roberta-large-nli-stsb-mean-tokens', 'distilbert-base-nli-stsb-mean-tokens', 'bert-base-nli-mean-tokens', 'bert-large-nli-mean-tokens', 'roberta-base-nli-mean-tokens', 'roberta-large-nli-mean-tokens', and 'distilbert-base-nli-mean-tokens'. 
The embedding vectors are then compared using cosine similarity, which produces a score between -1 and 1.
The rationale behind switching the applications at various stages, is to cache the results and create checkpoints for restarting any failed stages, easily. Additionally, since python provides better support for easy generation of embedding vectors, it was thus preferred over the Java based implementation, which is otherwise very beneficial for other tools. These applications were executed on a workstation running Ubuntu 20.04.2 LTS on top of AMD Ryzen 3 2200G, and 32GB ram.